
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                               %
% Chapter: BeamPositionMonitoring Thu Sep 24 16:02:10 EDT 2015
%                                                               %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Beam Position Monitoring}
\label{ch:BeamPositionMonitoring}

The Beam Position Monitors (BPMs) have played an important role in this analysis, and in
past analyses. They are the one means by which we can directly measure the approximate
transverse location of the beams. There are two sets of BPMs on either side of the PHENIX
IR, which we call "sector 7" and "sector 8". From these measurements, we can reconstrcut
the separation of the blue and yellow beams during a vernier scan. Our purpose in doing
so, is to correlate the beam displacement, to the event rate measured by the BBCs, in
order to directly measure the width of the beams in the x and y directions (assuming the
PHENIX standard coordinate system) \textcolor{red}{\textbf{missing-fig phenix coordinate
system}}. CAD executes these beam displacements through using DX magnets which deflect the
beams. The beams follow a planned presciption of movement, and are stepped into, and out
of alignment while PHENIX records data.

Past analyses have assumed that the data from the BPMS are unreliable. At times, the
polarity of the BPM measurement is flipped from the polarity of the planned step.
Regardless, we can use the BPM measurement to obtain a set of beam displacements, and
correlate those displacements with time, and with the BBC Rates using PRDF data, so the
utility is not insignificant. We can also identify which beam is scanned first from the
BPM data and the timing of each step. We can simply substitute planned step displacements
if we don't trust the value of beam displacements. We can see summarized the difference
betweeen planned steps, and real steps summarized in tables
\textcolor{red}{\textbf{missing-tab plannedsteps}}

\input{../BeamPositionMonitoring/tables/plannedsteps359711.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps360879.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps362492.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps364636.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps365866.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps366605.tex}
\input{../BeamPositionMonitoring/tables/plannedsteps367138.tex}

